Symbolic Weighted Language Models and\\ Parsing over Infinite Alphabets

### 2021-05-17 main.tex
cleanup

The \emph{summary} of a $\SWT$, resp. a $\SWA$, is ***

@@

% gen. to transducers
\begin{proposition}
Given a \SWA $A$ over alphabet $\Sigma$ 
and a commutative and bounded semiring $\Semiring$, 
there exists an effectively constructible \SWA $A'$ 
without $\epsilon$-transitions 
such that for all $s \in \Sigma^*$, $A'(s) = A(s)$.
\end{proposition}
%
\begin{proof}  ** Tb revised **   
Let $A = \< Q, \init, \<\wei_\epsilon, \wei_\Sigma>, \final >$.
The automaton $A'$ is $\< Q, \init, \<\wei'_\epsilon, \wei'_\Sigma>, \final >$,
where, for all $q, q' \in Q$,
$\wei'_\epsilon(q, q') = \zero$  
and 
% \[
% \begin{array}{rcccl}
% \wei'_\Sigma(q, q') & = &  & \displaystyle\bigoplus_{q_1 \in Q} &
%                \wei_\Sigma(q, q_1) \otimes  \wei_\epsilon(q_1, q')\\
%                     &   & \oplus  & \displaystyle\bigoplus_{q_2 \in Q} &
%                \wei_\epsilon(q, q_2) \otimes \wei_\Sigma(q_2, q')\\
% \end{array}
% \]
\[
\wei'_\Sigma(q, q') = 
\displaystyle\bigoplus_{q'' \in Q} \wei_\epsilon(q, q'') \otimes \wei_\Sigma(q'', q')
\oplus \displaystyle\bigoplus_{q_1 \in Q} \wei_\Sigma(q, q'') \otimes \wei_\epsilon(q'', q')
\]
In the above definition of $\wei'_\Sigma$ we use the operator
of product of function of $\Phi_\Sigma$ by $\Semiring$ 
described in Section~\ref{section:symbols}.
%
By definition of $\weight_{A}$ and distributivity of $\otimes$ on $\oplus$, 
** NO. TBC see \cite{Mohri02ijfcs} **
it holds that $\weight_{A}(q, s, q') = \weight_{A'}(q, s, q')$.
\qed
\end{proof}

@@

\begin{example}
structured words...
intro language of music notation ?
\end{example}

@@

\subsection{Best-first Search} 
\label{sec:best}\label{sec:search}
**hypotheses**
Let us assume that the semiring~$\Semiring$ is
commutative, bounded, complete, and total.
and 
** assume label theory is s.t. 
for all $\phi \in \Phii$, 
$\psi \in \Phir$, 
and $\eta \in \Phicr$, 
$\displaystyle\bigoplus_{a \in \Omegai} \phi(a)$ 
$\displaystyle\bigoplus_{r \in \Omegar} \phi(r)$ and 
$\displaystyle\bigoplus_{{\call{c}} \in \Omegac} 
\displaystyle\bigoplus_{{\return{r}} \in \Omegar} \eta({\call{c}}, {\return{r}})$
are computable...

@@

\noindent
The infinite sums in the updates of $d$ in Algorithm~\ref{algo:Dijkstra} are well defined
since~$\Semiring$ is complete.
The algorithm performs $2.|Q|^2$ iterations until $P$ is empty, 
and each iteration has a time complexity $O(|Q|^2 . |P|)$.
This gives a time complexity $O(|Q|^4 . |P|)$. 
It can be reduced by implementing $P$ as a priority queue, 
prioritized by the value returned by $d$
***complete***. %$|Q|^3.\log(|Q|^2)$

@@

\begin{proposition}
The problem of Symbolic Weighted  parsing 
can be solved in PTIME in the size of the input \SWT, \SWVPA (or \SWTA) and input word, 
and the computation time of the functions of the label theory.
\end{proposition}
%
\begin{proof}
Bar-Hillel construction
\qed  
\end{proof}

\subsection{Application to Automated Music Transcription}
\label{sec:transcription}
Symbolic Automated Music Transcription
and analysis of music performances

\subsubsection{Time Scales}
Real-Time Unit (RTU) = seconds

\noindent 
Musical-Time Unit (MTU) = number of measures

\noindent 
conversion via tempo value

\subsubsection{Representation of Music Performances}
We consider symbolic representations of musical performances, as finite sequences of events.
It corresponds to the concrete case of a MIDI file~\cite{SMF} 
recorded  from an electronic keyboard, 
or the output of a transcription from audio files~\cite{Benetos18AMTsurvey}.
%
For the sake of simplicity, 
we shall only consider here the case of monophonic performances, 
where at most one note is sounding at a time. 
The approach however extends to the polyphonic case.

A music performance is a finite sequence of events in a set~$\Sigma$.
Every event $e \in \Sigma$ has attributes such from a finite domain, 
like a number of key for a note 
or a flag indicating that it is a rest 
(\textsf{ON} and \textsf{OFF} messages in~\cite{SMF})
and a velocity value (0..127 in~\cite{SMF})).
%This representation is similar to the piano roll ~\cite{Muller15fundamentals} chap.1. 
Moreover, it contains a RTU value $\ioi{e}$ (real number) 
representing the time distance to the next event, 
or to the end of performance for the last event,
also called \emph{inter-onset interval}.

\subsubsection{Representation of Music Scores}
Music score are represented as structured words
made of timed %quantified 
events and parenthesized markups,
akin of nested words~\cite{AlurMadhusudan09nested}.

We consider an alphabet $\Delta$, every symbol of which is 
composed of a tag, in a finite set $\Xi$, 
and an MTU (rational) IOI duration value.
%The alphabet $\Delta$ 
It is partitioned into 
$\Delta = \Deltai \uplus \Deltac \uplus \Deltar$, 
like in Section~\ref{section:SWVPA}.
%
\noindent
The symbols of $\Deltai$ represent events:
% (infinite alphabet of internal symbols) made of:
with tags indicating a new note or grace-note (with null IOI), 
a rest or the continuation of the previous note (tie or dot).
%
The elements of $\Deltac \uplus \Deltar$ are matched
markups for describing the structure of the score, 
\ie the hierarchical grouping of events, and also, 
importantly the division of time in measures, tuplets...
%- parentheses for time divisions : tuplets, bars...
(linearization of rhythm trees \cite{jacquemard:hal-01138642}...).
They contain additional info such as tuple number, beaming policy...

\noindent
The duration values of letters of $\Delta$, in MTU (rational), 
can be computed with the markups and tags (\eg grace note has duration 0).

%\noindent
%There are simultaneous events, since grace notes has duration 0. They are ordered.
%
%\noindent
%Finite bound on the number of duration ratio. ?

\begin{example}
...      
\end{example}

\subsubsection{Performance/Score Distance Computation}
\label{app:distance}
We define a distance between performance and score representations
by a $\SWT$ $T = \< Q, \init, \wei, \final >$, over a semiring $\Semiring$.
** detail the elements of $\Semiring$ ....**
%are quadruplets of the form
%$\< t, s, \delta_t, \delta_s>$

Every state of $Q$ contains a 
tempo value in a finite domain (e.g. 30..300 bpm).
This value can be fixed 
or recomputed by the $T$ %transducer 
after reading each event, 
according to a perceptive/cognitive model of tempo 
such as~\cite{LargeJones99tempo}
(also used in the context of score following~\cite{Cont10TPAMI}).
% we wont detail here.

\subsection{Transcription by SW Parsing} %Best-first Search}
We assume a score language defined by a \SWVPA over the semiring 
$\Semiring$ of Section~\ref{app:distance}.

>>>>

%The \emph{summary} of a $\SWT$, resp. a $\SWA$, is ***

@@

% gen. to transducers
\begin{proposition}
Given a \SWA $A$ over alphabet $\Sigma$ 
and a commutative and bounded semiring $\Semiring$, 
there exists an effectively constructible \SWA $A'$ 
without $\epsilon$-transitions 
such that for all $s \in \Sigma^*$, $A'(s) = A(s)$.
\end{proposition}
%
%\begin{proof}  ** Tb revised **   
%Let $A = \< Q, \init, \<\wei_\epsilon, \wei_\Sigma>, \final >$.
%The automaton $A'$ is $\< Q, \init, \<\wei'_\epsilon, \wei'_\Sigma>, \final >$,
%where, for all $q, q' \in Q$,
%$\wei'_\epsilon(q, q') = \zero$  
%and 
%% \[
%% \begin{array}{rcccl}
%% \wei'_\Sigma(q, q') & = &  & \displaystyle\bigoplus_{q_1 \in Q} &
%%                \wei_\Sigma(q, q_1) \otimes  \wei_\epsilon(q_1, q')\\
%%                     &   & \oplus  & \displaystyle\bigoplus_{q_2 \in Q} &
%%                \wei_\epsilon(q, q_2) \otimes \wei_\Sigma(q_2, q')\\
%% \end{array}
%% \]
%\[
%\wei'_\Sigma(q, q') = 
%\displaystyle\bigoplus_{q'' \in Q} \wei_\epsilon(q, q'') \otimes \wei_\Sigma(q'', q')
%\oplus \displaystyle\bigoplus_{q_1 \in Q} \wei_\Sigma(q, q'') \otimes \wei_\epsilon(q'', q')
%\]
%In the above definition of $\wei'_\Sigma$ we use the operator
%of product of function of $\Phi_\Sigma$ by $\Semiring$ 
%described in Section~\ref{section:symbols}.
%%
%By definition of $\weight_{A}$ and distributivity of $\otimes$ on $\oplus$, 
%** NO. TBC see \cite{Mohri02ijfcs} **
%it holds that $\weight_{A}(q, s, q') = \weight_{A'}(q, s, q')$.
%\qed
%\end{proof}

@@

%\begin{example}
%structured words...
%intro language of music notation ?
%\end{example}

@@

\subsection{Best-first Search} 
\label{sec:best}\label{sec:search}
%**hypotheses**
Let us assume that the semiring~$\Semiring$ is
commutative, bounded, complete, and total.
and assume an effective label theory.
%s.t. for all $\phi \in \Phii$, 
%$\psi \in \Phir$, 
%and $\eta \in \Phicr$, 
%$\displaystyle\bigoplus_{a \in \Omegai} \phi(a)$ 
%$\displaystyle\bigoplus_{r \in \Omegar} \phi(r)$ and 
%$\displaystyle\bigoplus_{{\call{c}} \in \Omegac} 
%\displaystyle\bigoplus_{{\return{r}} \in \Omegar} \eta({\call{c}}, {\return{r}})$
%are computable...

@@

\noindent
The infinite sums in the updates of $d$ in Algorithm~\ref{algo:Dijkstra} are well defined
since~$\Semiring$ is complete.
The algorithm performs $2.|Q|^2$ iterations until $P$ is empty, 
and each iteration has a time complexity $O(|Q|^2 . |P|)$.
This gives a time complexity $O(|Q|^4 . |P|)$. 
%
%It can be reduced by implementing $P$ as a priority queue, 
%prioritized by the value returned by $d$
%***complete***. %$|Q|^3.\log(|Q|^2)$

@@

\begin{proposition}
The problem of Symbolic Weighted  parsing 
can be solved in PTIME in the size of the input \SWT, \SWVPA (or \SWTA) and input word, 
and the computation time of the functions of the label theory.
\end{proposition}
%
%\begin{proof}
%Bar-Hillel construction
%\qed  
%\end{proof}
